 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Large Colored Title Article
% LaTeX Template
% Version 1.1 (25/11/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Code Snippet
% LaTeX Template
% Version 1.0 (14/2/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Velimir Gayevskiy (vel@latextemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[11pt,oneside]{book}
%\documentclass[DIV=calc, paper=a4, fontsize=11pt, twocolumn]{scrartcl}	 % A4 paper and 11pt font size
%----------------------------------------------------------------------------------------

\usepackage{listings} % Required for inserting code snippets
\usepackage[usenames,dvipsnames]{color} % Required for specifying custom colors and referring to colors by name
\usepackage{tocbibind}

%%%%%%%%%% Uncomment if you need to write some code
% \definecolor{DarkGreen}{rgb}{0.0,0.4,0.0} % Comment color
% \definecolor{highlight}{RGB}{255,251,204} % Code highlight color
%
% \lstdefinestyle{Style1}{ % Define a style for your code snippet, multiple definitions can be made if, for example, you wish to insert multiple code snippets using different programming languages into one document
% language=Java, % Detects keywords, comments, strings, functions, etc for the language specified
% backgroundcolor=\color{highlight}, % Set the background color for the snippet - useful for highlighting
% basicstyle=\footnotesize\ttfamily, % The default font size and style of the code
% breakatwhitespace=false, % If true, only allows line breaks at white space
% breaklines=true, % Automatic line breaking (prevents code from protruding outside the box)
% captionpos=b, % Sets the caption position: b for bottom; t for top
% commentstyle=\usefont{T1}{pcr}{m}{sl}\color{DarkGreen}, % Style of comments within the code - dark green courier font
% deletekeywords={}, % If you want to delete any keywords from the current language separate them by commas
% %escapeinside={\%}, % This allows you to escape to LaTeX using the character in the bracket
% firstnumber=1, % Line numbers begin at line 1
% frame=single, % Frame around the code box, value can be: none, leftline, topline, bottomline, lines, single, shadowbox
% frameround=tttt, % Rounds the corners of the frame for the top left, top right, bottom left and bottom right positions
% keywordstyle=\color{Blue}\bf, % Functions are bold and blue
% morekeywords={}, % Add any functions no included by default here separated by commas
% numbers=left, % Location of line numbers, can take the values of: none, left, right
% numbersep=10pt, % Distance of line numbers from the code box
% numberstyle=\tiny\color{Gray}, % Style used for line numbers
% rulecolor=\color{black}, % Frame border color
% showstringspaces=false, % Don't put marks in string spaces
% showtabs=false, % Display tabs in the code as lines
% stepnumber=5, % The step distance between line numbers, i.e. how often will lines be numbered
% stringstyle=\color{Purple}, % Strings are purple
% tabsize=4, % Number of spaces per tab in the code
% }
%
% % Create a command to cleanly insert a snippet with the style above anywhere in the document
% \newcommand{\insertcode}[2]{\begin{itemize}\item[]\lstinputlisting[caption=#2,label=#1,style=Style1]{#1}\end{itemize}} % The first argument is the script location/filename and the second is a caption for the listing
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[utf8x]{inputenc}
\usepackage{lipsum} % Used for inserting dummy 'Lorem ipsum' text into the template
\usepackage[english,italian]{babel} % English language/hyphenation
\usepackage[protrusion=true,expansion=true]{microtype} % Better typography
\usepackage{amsmath,amsfonts,amsthm} % Math packages
\usepackage{amssymb} % AMS symbols
\usepackage[lined,boxed,commentsnumbered]{algorithm2e}
\usepackage[svgnames]{xcolor} % Enabling colors by their 'svgnames'
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{fix-cm}	 % Custom font sizes - used for the initial letter in the document
\usepackage{graphicx}
\usepackage{adjustbox} % resize only if the table goes beyond \textwidth, otherwise not
\usepackage{sectsty} % Enables custom section titles
\allsectionsfont{\usefont{OT1}{phv}{b}{n}} % Change the font of all section commands

%----------------------------------------------------------------------------------------
%	Clickable Links
\usepackage{hyperref}
\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=black,
    urlcolor=black
}
%----------------------------------------------------------------------------------------

\usepackage{fancyhdr} % Needed to define custom headers/footers
\pagestyle{fancy} % Enables the custom headers/footers
\usepackage{lastpage} % Used to determine the number of pages in the document (for "Page X of Total")
\usepackage{graphicx}
\usepackage{subfig}
% Headers - all currently empty
\lhead{}
\chead{}
\rhead{}

% Footers
\lfoot{}
\cfoot{}
\rfoot{\footnotesize Pagina \thepage\ di \pageref{LastPage}} % "Pagina 1 di 2"

\renewcommand{\headrulewidth}{0.0pt} % No header rule
\renewcommand{\footrulewidth}{0.4pt} % Thin footer rule

\usepackage{lettrine} % Package to accentuate the first letter of the text
\newcommand{\initial}[1]{ % Defines the command and style for the first letter
\lettrine[lines=3,lhang=0.3,nindent=0em]{
\color{DarkGoldenrod}
{\textsf{#1}}}{}}

\newcommand{\ggll}{\mathrel{\substack{\ll\\[-.05em]\gg}}} % convergence symbol
\SetKwProg{Fn}{Function}{}{} % Function keyword in algorithm definition

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\usepackage{titling} % Allows custom title configuration

\newcommand{\HorRule}{\color{DarkGoldenrod} \rule{\linewidth}{1pt}} % Defines the gold horizontal rule around the title

\pretitle{\vspace{-30pt} \begin{flushleft} \HorRule \fontsize{30}{30} \usefont{OT1}{phv}{b}{n} \color{DarkRed} \selectfont} % Horizontal rule before the title

\title{Smart Grid} % Your article title

\posttitle{\par\end{flushleft}\vskip 0.5em} % Whitespace under the title

\preauthor{\begin{flushleft}\large \lineskip 0.5em \usefont{OT1}{phv}{b}{sl} \color{DarkRed}} % Author font configuration

\author{Marco Amoruso, Daniele Anello, Francesco Farina, \\Iolanda Rinaldi } % Your name

\postauthor{\footnotesize \usefont{OT1}{phv}{m}{sl} \color{Black} % Configuration for the institution name
\newline
\newline
Università degli Studi di Salerno % Your institution

\par\end{flushleft}\HorRule} % Horizontal rule after the title

\date{Last edit: \today} % Add a date here if you would like one to appear underneath the title block

%----------------------------------------------------------------------------------------

\begin{document}

\maketitle % Print the title

\thispagestyle{fancy} % Enabling the custom headers/footers for the first page

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

% The first character should be within \initial{}

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------
\tableofcontents
\chapter{Introduzione}
\input{Intro.tex}
\chapter{Smart Grid: panorama attuale e definizione}
\input{Capitolo1.tex}
\chapter{Architettura}
\input{Capitolo2.tex}
\chapter{Smart Grid Cybersecurity}
%\chapter{Smart Grid Cybersecurity: requisiti, building blocks e potenziali minacce}
\input{Capitolo3.tex}
\chapter{Standard e tecnologie \label{chap:chap5}}
\input{Capitolo4.tex}
\chapter{Principali vulnerabilità delle Smart Grid: attacchi e contromisure}
\input{Capitolo5.tex}
%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------

\begin{thebibliography}{99}
\bibitem{securingSG} Tony Flick and Justin Morehouse. \emph{Securing the Smart Grid. Next Generation Power Grid Security.} Elsevier Inc, 2011.
\bibitem{smartgrids} Stuart Borlase. \emph{Smart Grids: Infrastructure, Technology, and Solutions.} CRC Press, 2013.
\bibitem{smartgrid} Janaka Ekanayake, Kithsiri Liyanage, Jianzhong Wu, Akihiko Yokoyama and Nick Jenkins. \emph{Smart Grid: Technology and Applications.} Wiley, 2012.
\bibitem{pathsg} Hassan Farhangi. The Path of the Smart Grid. In \emph{IEEE power \& energy magazine}, pages 18-28, january/february 2010.
\bibitem{surveymotiv} Ye Yan, Yi Qian, Hamid Sharif, and David Tipper. A Survey on Smart Grid Communication Infrastructures: Motivations, Requirements and Challenges. In \emph{IEEE communications surveys \& tutorials,} vol.15, no.1, 2013.
\bibitem{surveyrouting} Nico Saputro, Kemal Akkaya and Suleyman Uludag. \emph{A survey of routing protocols for smart grid communications}. In Computer Networks 56, pages 2742–2771, 2012.
\bibitem{improvedgrid} Xi Fang, Satyajayant Misra, Guoliang Xue and Dejun Yang. \emph{Smart Grid – The New and Improved Power Grid: A Survey}. In IEEE communications surveys \& tutorials, vol.14, no.4, 2012.
\bibitem{topologies} Fundamental of Electricity: Radial, Loop, \& Network Systems. \url{http://epb.apogee.net/foe/ftdstr.asp}
\bibitem{nist} NIST Website about Smart Grids. \url{http://www.nist.gov/smartgrid/}
\bibitem{gungor} Gungor VC, et al. \emph{A survey on smart grid potential applications and
communication requirements}. IEEE Trans Ind Inform. 2013;9(1):28–42.
\bibitem{smartmet} M. Lee, O. Aslam, B. Foster, D. Kathan, J. Kwok, L. Medearis,
R. Palmer, P. Sporborg, and M. Tita. \emph{Assessment of demand response
and advanced metering}. Federal Energy Regulatory Commission, Tech. Rep., 2013 \url{https://www.ferc.gov/legal/staff-reports/12-20-12-demand-response.pdf}
\bibitem{energycons} C. McKerracher, J. Torriti. \emph{Energy consumption feedback in perspective: integrating australian data to meta-analyses on in home displays}. Energy Effic. 2012;6(2), 387–405.
\bibitem{vehicl} Morgan Trevor. \emph{Smart Grids and Electric Vehicles Made for Each Other}. Germany: International Transport Forum on Seamless Transport, Discussion Paper; 2012.
\bibitem{visionfr} F. Li, W. Qiao, H. Sun, H. Wan, J. Wang, Y. Xia, Z. Xu, and P. Zhang.
\emph{Smart transmission grid: Vision and framework}. IEEE Trans. on Smart Grid, 1(2):168–177, 2010.
\bibitem{voltsta} K. Vu, M. M. Begovic, D. Novosel, and M. M. Saha. \emph{Use of local measurements to estimate voltage-stability margin} IEEE Trans. Power Syst., vol. 14, no. 3, pp. 1029–1035, Aug. 1999. 
\bibitem{stavolt} B. Milosevic and M. Begovic. \emph{Voltage-stability protection and control using a wide-area network of phasor measurements} IEEE Trans. Power Syst., vol. 18, no. 1, pp. 121–127, Feb. 2003. 
\bibitem{tras1} D. Povh, D. Retzmann, J. Kreusel. \emph{Integrated AC/DC transmission systems — Benefits of power electronics for security and sustainability of power supply}. PSCC 2008, Glasgow, U.K., July 14–17, 2008. 
\bibitem{tras2} European Technology Platform. \emph{Vision and Strategy for Europe’s Electricity Networks of the Future}. European Commission. \url{https://ec.europa.eu/research/energy/pdf/smartgrids_en.pdf}
\bibitem{tras3} W. Breuer, D. Povh, D. Retzmann, and E. Teltsch. \emph{Trends for future HVDC Applications}. 16th CEPSI, Mumbai November 6–10, 2006.
\bibitem{dist1} Yvonne-Anne Pignolet, Holger Elias, Timo Kyntäjä, Ignacio Martín Díaz de Cerio, Jürgen Heiles, Didier Boëda, Raphael Caire. \emph{Future Internet for Smart Distribution Systems}. IEEE PES Innovative Smart Grid Technologies (ISGT) Europe Conference, 2012.
\bibitem{802.3} Low, D. (2011) IEEE 802.3 Ethernet, IEEE, March 2011, \url{http://www.ieee802.org/minutes/2011-March/802\%20workshop/index.shtml}.
\bibitem{zb} ZigBee. \url{http://www.zigbee.org}
\bibitem{wimax} Cudak, M. (ed.) (2010) IEEE 802.16m System Requirements, IEEE 802.16 Task Group M, January 2010, \url{http://ieee802.org/16/tgm/docs/80216m-07_002r10.pdf}.
\bibitem{dnp} IEEE Power \& Energy Society (2010) IEEE Std 1815TM-2010. \emph{IEEE Standard for Electric Power Systems Communications-Distributed Network Protocol (DNP3)}, June.
\bibitem{iec61850} Jianqing Zhangand Carl A. Gunter. \emph{IEC 61850 - communication networks and systems in substation: an overview of computer science}.
University of Illinois at Urbana Champaign, Jan. 2010.
\bibitem{tls} T. Dierks, E. Rescorla. \emph{RFC 5246: The Transport Layer Security (TLS) Protocol}. Version 1.2, August 2008.
\bibitem{standard14908} International Organization for Standardization. \emph{ISO/IEC 14908-1:2012: Information technology – Control network protocol – Part 1: Protocol stack}, 2012.
\bibitem{cbcpadding} Vaudenay, S.: \emph{Security Flaws Induced by CBC Padding — Applications to SSL, IPSEC, WTLS ... }In: Knudsen, L.R. (ed.) Advances in Cryptology — EUROCRYPT 2002, International Conference on the Theory and Applications of Cryptographic Techniques, Amsterdam, The Netherlands, April 28 – May 2, 2002, Proceedings. Lecture Notes in Computer Science, vol. 2332, pp. 534–546. Springer (2002)
\bibitem{osgp} ETSI: Open Smart Grid Protocol (OSGP). Reference DGS/OSG-001, European Telecommunications Standards Institute, Sophia Antipolis Cedex – France (January 2012). \url{http://www.osgp.org/}
\bibitem{dumbcrypto} Jovanovic, Philipp, and Samuel Neves. \emph{Dumb Crypto in Smart Grids: Practical Cryptanalysis of the Open Smart Grid Protocol}.
\bibitem{wep1} Andreas Klein. \emph{Attacks on the RC4 stream cipher}. Des. Codes Cryptography, 48(3):269–286, 2008.
\bibitem{wep2} Erik Tews, Ralf-Philipp Weinmann, and Andrei Pyshkin. \emph{Breaking 104 bit WEP in less than 60 seconds.} In Sehun Kim, Moti Yung, and Hyung-Woo Lee, editors, Information Security Applications, 8th International Workshop, WISA 2007, Jeju Island, Korea, August 27-29, 2007, Revised Selected Papers, volume 4867 of Lecture Notes in Computer Science, pages 188–202. Springer, 2007.
\bibitem{rc4tls} AlFardan, N.J., Bernstein, D.J., Paterson, K.G., Poettering, B., Schuldt, J.C.N.: \emph{On the Security of RC4 in TLS}. In: King, S.T. (ed.) Proceedings of the 22th USENIX Security Symposium, Washington, DC, USA, August 14–16, 2013. pp. 305–320. USENIX Association (2013).
\bibitem{weakrc4} Fluhrer, S.R., Mantin, I., Shamir, A.: \emph{Weaknesses in the Key Scheduling Algorithm of RC4}. In: Vaudenay, S., Youssef, A.M. (eds.) Selected Areas in Cryptography, 8th Annual International Workshop, SAC 2001 Toronto, Ontario, Canada, August 16-17, 2001, Revised Papers. Lecture Notes in Computer Science, vol. 2259, pp. 1–24. Springer (2001)
\bibitem{rc4statistical} Fluhrer, S.R., McGrew, D.A.: \emph{Statistical Analysis of the Alleged RC4 Keystream Generator}. In: Schneier, B. (ed.) Fast Software Encryption, 7th International Workshop, FSE 2000, New York, NY, USA, April 10–12, 2000, Proceedings. Lecture Notes in Computer Science, vol. 1978, pp. 19–30. Springer (2000)
\bibitem{rc4rand} Gupta, S.S., Maitra, S., Paul, G., Sarkar, S.: \emph{(Non-)Random Sequences from (Non-)Random Permutations – Analysis of RC4 Stream Cipher}. J. Cryptology 27(1), 67–108 (2014).
\bibitem{aircrackng} Aircrack-ng. \url{http://www.aircrack-ng.org/}
\bibitem{kali} Kali Linux. \url{https://www.kali.org/}
\bibitem{wireshark} Wireshark. \url{https://www.wireshark.org/}
\bibitem{nmap} Nmap. \url{https://www.nmap.org/}
\bibitem{nessus} Nessus. \url{https://www.nessus.org/}
\bibitem{cainabel} Cain \& Abel. \url{http://www.oxid.it/}
\bibitem{metasploit} Metasploit. \url{http://www.metasploit.com/}
\bibitem{monticelli} A. Monticelli. \emph{Electric Power System State Estimation}. Proc. IEEE, vol. 88, Feb. 2000, pp. 262–82.
\bibitem{baddatainj} M. Esmalifalak, Z. Han, and L. Song, \emph{Effect of Stealthy Bad Data Injection On Network Congestion In Market Based Power System}. IEEE WCNC 2012, Paris, France, Apr. 2010.
\bibitem{falsedatainj} Y. Liu, M. K. Reiter, and P. Ning. \emph{False Data Injection Attacks Against State Estimation in Electric Power Grids}. 16th ACM Conf. Computer and Commun. Security, Gaithersburg, MD, Nov. 2009, pp. 21–30.
\bibitem{baddatainj2} L. Xie, Y. Mo, and B. Sinopoli. \emph{False Data Injection Attacks in Electricity Markets}. 1st IEEE Int’l. Conf. Smart Grid Commun., Gaithersburg, MD, Oct. 2010, pp. 226–31.
\bibitem{baddatainjattackdef} Huang, Yi, et al. \emph{Bad data injection in smart grid: attack and defense mechanisms}. Communications Magazine, IEEE 51.1 (2013): 27-33.
\bibitem{convdef1} A. Abur and A. G. Exposito. \emph{Power System State Estimation: Theory and Implementation}. Marcel Dekker, 2004.
\bibitem{convdef2} A. J. Wood and B. F. Wollenberg. \emph{Power Generation, Operation, and Control}. Wiley, 1996.
\bibitem{lorden} H. V. Poor and Q. Hadjiliadis. \emph{Quickest Detection}. Cambridge Univ. Press, 2008.
\bibitem{esmalifalak} Esmalifalak, M., Huy Nguyen, Rong Zheng, Zhu Han.\emph{Stealth false data injection using independent component analysis in smart grid} Smart Grid Communications (SmartGridComm) 2011 IEEE International Conference on , vol., no., pp.244-248, 17-20 Oct. 2011.
\bibitem{lica} J. Himberg and A. Hyvarinen, \emph{Independent Component Analysis for Binary Data: An Experimental Study}. 3rd Int’l. Conf. Independent Component Analysis and Blind Signal Separation, Malm, Sweden, June 2001.
\bibitem{offswitch} R. Anderson and S. Fuloria, \emph{Who controls the off switch?}. Proc. of the Conference on Smart Grid Communications (SmartGridComm), 2010.
\bibitem{remotecontrol} M. Costache, V. Tudor, M. Almgren, M. Papatriantafilou and C. Saunders. \emph{Remote control of smart meters: friend or foe?}. Proc. of the European Conference on Computer Network Defense (EC2ND), 2011.
\bibitem{amithreats} D. Grochocki, J. H. Huh, R. Bobba, W. H. Sanders, A. A. Cardenas, and J. G. Jetcheva, \emph{AMI threats, intrusion detection requirements and deployment recommendations}. Proc. of the Conference on Smart Grid Communications (SmartGridComm), 2012.
\bibitem{disconnectfaqs} \emph{GE I-210+ remote disconnect FAQs} \url{http://www.tantalus.com/pdf/products/TUNet_TC_1110-1210RD_plus-R3-2.pdf}.
\bibitem{delay} Temple, William G., Binbin Chen and Nils Ole Tippenhauer. \emph{Delay makes a difference: Smart grid resilience under remote meter disconnect attack}. Smart Grid Communications (SmartGridComm). IEEE International Conference on. IEEE, 2013.
\bibitem{radio} A. Ghassemi, S. Bavarian, and L. Lampe, \emph{Cognitive radio for smart grid communications}. Proc. of the Conference on Smart Grid Communications (SmartGridComm), 2010.
\bibitem{toggle1} \emph{Avista utilities update to idaho public utility commission staff on remote reconnect/disconnect pilot} \url{http://www.puc.idaho.gov/fileroom/cases/elec/AVU/AVUE0709/company/20130211UPDATE\%20ON\%20REMOTE\%20RECONNECT\%20DISCONNECT.PDF}.
\bibitem{toggle2} \emph{Texas-new mexico power company’s request for approval of an advanced metering system (ams) deployment and ams surcharge} \url{https://www.smartgrid.gov/files/TexasNew_Mexico_Power_Company_Request_For_Approval_Advance_201005.pdf}.
\bibitem{demand} W. H. Kersting. \emph{Distribution system modeling and analysis}. CRC Press, LLC, 2012.
\bibitem{jam} F. Li and R. Bo. \emph{Congestion and price prediction under load variation}. IEEE Control System Magazine, vol.19, pp.59–70, Oct. 2009.
\bibitem{diff} Lipmaa, H., Moriai, S.: E cient Algorithms for Computing Differential Properties of Addition. In: Matsui, M. (ed.) Fast Software Encryption, 8th International Workshop, FSE 2001 Yokohama, Japan, April 2–4, 2001 Revised Papers. Lecture Notes in Computer Science, vol. 2355, pp. 336–350. Springer (2001).
\bibitem{jamming} Li, Husheng, and Zhu Han. ``Manipulating the electricity power market via jamming the price signaling in smart grid.'' GLOBECOM Workshops (GC Wkshps), 2011 IEEE. IEEE, 2011.
\end{thebibliography}
%----------------------------------------------------------------------------------------

\end{document}